---
title: "Lab 5 - Task 1"
author: "Anna Palmqvist Sjövall"
date: "30 augusti 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=15)
```

## 2. Load house_data from you working directory (folder). Consider "sqft_living" as X and "price" as Y

```{r}
# upload data
data <- read.csv("house_data.csv", header=TRUE, stringsAsFactors=FALSE)
X <- data$sqft_living
Y <- data$price
head(data)
```

# 3. Plot X versus Y: plot (x, y).
```{r}
plot(X,Y)
```

# 4. Calculate average of X and average of Y: mean (X), mean (Y)
```{r}
X_avg <- mean(X)
Y_avg <- mean(Y)
X_avg
Y_avg
```

#5/6. Calculate X - mean (X) and Y - mean (Y)
```{r}
X_fac <- X - X_avg
data["X - mean(X)"] <- X_fac
Y_fac <- Y - Y_avg
data["Y - mean(Y)"] <- Y_fac
head(data)
```

#7. Calculate (X - mean (X)) * (Y - mean (Y))
```{r}
XY <-X_fac * Y_fac
data["(X - mean (X)) * (Y - mean (Y))"] <- XY
head(data)
```

#8. Calculate sum of ((X - mean (X)) * (Y - mean (Y)))
```{r}
XY_sum <- sum(XY)
XY_sum
```

#9. Calculate (X - mean (X)) ^2
```{r}
X_sq <- X_fac ^ 2
data["(X - mean (X)) ^2"] <- X_sq
head(data)
```

#10. Calculate sum of ((X - mean (X)) ^2)
```{r}
X_sum <- sum(X_sq)
X_sum
```

#11. Calculate ??1: ??1=Divide Step 8 by Step 10
```{r}
beta1 <- XY_sum / X_sum
beta1
```

#12. Calculate ??0: ??0= mean (Y) - ??1 * mean (X)
```{r}
beta0 <- Y_avg - beta1*X_avg
beta0
```

#13. Calculate Yest= ??0 + ??1 * X
```{r}
Y_est <- beta0 + beta1*X
head(Y_est)
```

#14. Display the Yest as follows
```{r}
table <- data.frame(sqft_living_X=X, price_Y=Y, Y_est=Y_est)
head(table)
```

#15. Calculate the error: Error= price Y- Yest
```{r}
table["Error= price Y - Y_est"] <- Y - Y_est
head(table)
```

#16. Calculate Root Mean Squared Error (RMSE):
```{r}
RMSE <- sqrt(sum((Y_est - Y)^2) / length(Y))
RMSE
```

#Display Result
```{r}
plot (X,Y)
lines(X, Y_est)
```
