---
title: "Lab 9"
author: "Anna Palmqvist Sjövall"
date: "3 oktober 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# upload data
```{r}
data <- read.csv("data1.csv", header=TRUE, stringsAsFactors=FALSE)
data
minSupp <- 2
```

# Step 1. Calculate the support/frequent of all items.
```{r}
c1 <- data.frame(colSums(data[-1])) #Ignore Transaction ID
colnames(c1) <- "SupportFrequent"
c1
```

# Step 2. Discard the items with the support less than the pre-defined threshold.
```{r}
c1 <- subset(c1, c1$SupportFrequent >= minSupp)
c1
```

# Step 3. Combine two items and calculate their support/frequent.
```{r}
names <- rownames(c1)
#Get all permutations
permutations <- (merge(names, names))
#Filter out when x==y and same permutations
c2 <- subset(permutations, permutations[1] != permutations[2] & duplicated(t(apply(permutations, 1, sort))))

#Takes a list of items and returns the number of times they are found together in data.
freqInData <- function(x) {
  df <- data.frame(matrix(ncol = length(x), nrow = dim(data)[1]))
  for(i in 1:length(x)){
    df[i] <- data[x[i]]
  }
  sum(Reduce('+', df) == length(x))
}

c2$SupportFrequent <- apply(c2, 1, freqInData)
c2

```

# Step 4. Discard the items with the support less than the pre-defined threshold.
```{r}
c2 <- subset(c2, c2$SupportFrequent >= minSupp)
c2
```


# Step 5. Combine three items and calculate their support/frequent.
```{r}
#We need the x value (could have also chosen y) to be equal to form the trio
equalX <- split(c2, c2$x)
dataFrames <- equalX[sapply(equalX, nrow) >= 2]
nbrDataFrames <- length(sapply(dataFrames, length))
c3 <- matrix(NA, nrow = nbrDataFrames, ncol = 3)

#Go through all possible trios and form them
for(i in 1:nbrDataFrames) {
  elemX <- toString(dataFrames[[i]]$x[1])
  elemY <- unlist(strsplit(toString(dataFrames[[i]]$y), ", "))
  c3[i,] <- c(elemX, elemY)
}

c3 <- data.frame(c3)
c3$SupportFrequent <- apply(c3, 1, freqInData)
c3

```

# Step 6. Discard the items with the support less than the pre-defined threshold.

```{r}
c3 <- subset(c3, c3$SupportFrequent >= minSupp)
c3
```

