---
title: "Lab 3"
author: "CSE5DEV"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Welcome to CSE5DEV**

You need to download the R code files and datasets into your working folder (current directory)


This lab will cover Lecture 2 and Lecture 3. 


## Task 2 - Lecture 3: Data Cleaning & Normalisation. 

We will perfrom the following stpes:

- Data Cleaning.
- Data Transformation
- Data Normalisation.

## Code chunks

## Load the dataset

We will look at dirty_iris.csv data

```{r}
# upload dirty_iris.csv data
dirty_iris <- read.csv("dirty_iris.csv", header=TRUE, stringsAsFactors=FALSE)
dirty_iris

```
## Call is.na() on the full dirty_iris to spot all NAs
```{r}
is.na(dirty_iris)
```
## Call is.na() on the full Species feature to spot all NAs


```{r}
is.na(dirty_iris$Species)
```
## Use the any() function to ask whether there are any NAs in the data
```{r}
any(is.na(dirty_iris))
```
## View a summary() of the dataset
```{r}
summary(dirty_iris)

```



## Replace all empty strings in status with NA
```{r}
dirty_iris$Sepal.Length[dirty_iris$Sepal.Length == ""] <- NA
dirty_iris$ Sepal.Width [dirty_iris$ Sepal.Width == ""] <- NA
dirty_iris$ Petal.Length[dirty_iris$ Petal.Length == ""] <- NA
dirty_iris$Petal.Width [dirty_iris$Petal.Width == ""] <- NA
dirty_iris$Species[dirty_iris$Species == ""] <- NA
summary(dirty_iris)
```
## Use complete.cases() to see which rows have missing values
```{r}
complete.cases(dirty_iris)
```


## Count missing values whole data
```{r}
sum(is.na(dirty_iris))
```

## Count missing values in a feature

```{r}
sum(is.na(dirty_iris$Sepal.Length))
```
# Find indices of NAs in Sepal.Length

```{r}
ind <- which(is.na(dirty_iris$Sepal.Length))
ind
```










## Look at the full rows for records missing Sepal.Length
```{r}
dirty_iris[ind, ]
```









# Set Sepal.Length missing values to 0.5
```{r}

dirty_iris$Sepal.Length[ind] <- 0.5
dirty_iris[ind, ]
```

# We can omit all rows with any missing values using na.omit() 
```{r}
na.omit(dirty_iris)
```

# Get index of row with -1 value

```{r}
m<-max(dirty_iris$Sepal.Length)
ind0 <- which(dirty_iris$Sepal.Length ==m ) 
print (ind0)
dirty_iris$Sepal.Length[ind0]
```





## Your turn
- Replace missing values in Sepal.Width by mean value
```{r}
#Calculate mean of the Sepal.Width values which are not NA
mean_sepal_width <- mean(na.omit(dirty_iris)$Sepal.Width)
mean_sepal_width
#Replace the NA values with the mean value
dirty_iris$Sepal.Width[ind] <- mean_sepal_width
#Display data which used to have missing values in Sepal.Width column
dirty_iris[ind, ]
```
- Replace Missing valuse in Petal.Length by max value
```{r}
#Index of missing values in Petal.Length
ind2 <- which(is.na(dirty_iris$Petal.Length))
#Calculate max of the Petal.Length values which are not NA
max_petal_length <- max(na.omit(dirty_iris)$Petal.Length)
max_petal_length
#Replace the NA values with the max value
dirty_iris$Petal.Length[ind2] <- max_petal_length
#Display data which used to have missing values in Petal.Length column
dirty_iris[ind2, ]
```
- Convert Species names into: setosa to "1", versicolor to "2" and virginica to "3"
```{r}
ind_setosa <- which(dirty_iris$Species == "setosa")
dirty_iris$Species[ind_setosa] <- "1"

ind_versicolor <- which(dirty_iris$Species == "versicolor")
dirty_iris$Species[ind_versicolor] <- "2"

ind_virginica <- which(dirty_iris$Species == "virginica")
dirty_iris$Species[ind_virginica] <- "3"

dirty_iris$Species
```
- Normalise Sepal.Width by Min-Max: Sepal.Width= Sepal.Width - min / max -min
```{r}
min_sepal_width <- min(na.omit(dirty_iris)$Sepal.Width)
min_sepal_width
max_sepal_width <- max(na.omit(dirty_iris)$Sepal.Width)
max_sepal_width

differance = max_sepal_width - min_sepal_width

dirty_iris$Normalized.Min.Max <- (dirty_iris$Sepal.Width - min_sepal_width)/differance

dirty_iris$Normalized.Min.Max
```
- Normalise Sepal.Width Z score: Sepal.Width= (Sepal.Width- mean)/std
```{r}
mean_sepal_width <- min(na.omit(dirty_iris)$Sepal.Width)
mean_sepal_width
std_sepal_width <- sd(na.omit(dirty_iris)$Sepal.Width)
std_sepal_width

dirty_iris$Normalized.Z <- (dirty_iris$Sepal.Width - mean_sepal_width)/std_sepal_width

dirty_iris$Normalized.Z
```

```{r}
dirty_iris
```

