---
title: "Lab 3"
author: "CSE5DEV"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





## **Welcome to CSE5DEV**

You need to download the R code files and datasets into your working folder (current directory)


This lab will cover Lecture 2 and Lecture 3. 



## Task 1- Lecture 2: Data exploration.

We will perfrom the following stpes:

-Summary and stats.

-Explore Individual Variables.

-Exploration of multiple variables.

## Code chunks

## Load the iris dataset

The "iris" dataset will be used for Data exploration. 

```{r}
# The datasets package needs to be loaded to access our data 
# For a full list of these datasets, type library(help = "datasets")
library(datasets)
data(iris)
```



## Or we can upload the dataset from the working folder (directory)

```{r}
mydata <- read.csv("iris.csv", header=TRUE)
head(mydata)
```

## Display the Size of Data
```{r}
dim(iris)
```
## Display features (columns) names
```{r}
names(iris)
```

## Calculate number of rows in a dataset
```{r}
nrow(iris) 
```

## Calculate number of columns in a dataset
```{r}
ncol(iris) 
```


## Display the structure Of data
```{r}
str(iris)
```

## Display attributes of data
```{r}
attributes(iris)
```

## Display the first three rows of data
```{r}
iris[1:3, ]
```

## Display head of data - first three rows
```{r}
head(iris, 3)
```
## Display tail of data - last three rows
```{r}
tail(iris, 3)
```

## Display a single feature (Column)
```{r}
#The first 10 values of Sepal.Length
iris[1:10, "Sepal.Length"]
#or
iris$Sepal.Length[1:10]
```

## Summary of Data
```{r}
summary(iris)
```
## Calculate range, minimum, maximum, mean, median, std and the firrst (25%) and third (75%) quartiles for numeric variables.
```{r}
range(iris$Sepal.Length)
min(iris$Sepal.Length)
max(iris$Sepal.Length)
mean(iris$Sepal.Length)
median(iris$Sepal.Length)
sd(iris$Sepal.Length)
quantile(iris$Sepal.Length)
```
## Calculate Variance
```{r}
var(iris$Sepal.Length)
```

## Calculate Frequency for categorical variables
```{r}
table(iris$Species)
```
## Calculate Covariance of two variables
```{r}
cov(iris$Sepal.Length, iris$Petal.Length)
```

## Calculate Correlation of two variables
```{r}
cor(iris$Sepal.Length, iris$Petal.Length)
```




## Plot Sepal.Length vs Sepal.Width
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width)
```
## Plot  Sepal.Length only
```{r}
plot(iris$Sepal.Length)
```



```{r}
##
```

## Your turn

Try creating a code chunk here to do the following steps:

- Perform all above steps for 1) Sepal.Width, 2) Petal.Length  and  3) Petal.Width 
-> 1) Sepal.Width
```{r}
#The first 10 values of Sepal.Width
iris[1:10, "Sepal.Width"]
```
```{r}
## Calculate range, minimum, maximum, mean, median, std and the firrst (25%) and third (75%) quartiles for numeric variables.
range(iris$Sepal.Width)
min(iris$Sepal.Width)
max(iris$Sepal.Width)
mean(iris$Sepal.Width)
median(iris$Sepal.Width)
sd(iris$Sepal.Width)
quantile(iris$Sepal.Width)
```
```{r}
## Calculate Variance
var(iris$Sepal.Width)
```
```{r}
## Plot  Sepal.Width only
plot(iris$Sepal.Width)
```


-> 2) Petal.Length
```{r}
#The first 10 values of Petal.Length
iris[1:10, "Petal.Length"]
```
```{r}
## Calculate range, minimum, maximum, mean, median, std and the firrst (25%) and third (75%) quartiles for numeric variables.
range(iris$Petal.Length)
min(iris$Petal.Length)
max(iris$Petal.Length)
mean(iris$Petal.Length)
median(iris$Petal.Length)
sd(iris$Petal.Length)
quantile(iris$Petal.Length)
```
```{r}
## Calculate Variance
var(iris$Petal.Length)
```
```{r}
## Plot Petal.Length only
plot(iris$Petal.Length)
```


-> 3) Petal.Width
```{r}
#The first 10 values of Petal.Width
iris[1:10, "Petal.Width"]
```
```{r}
## Calculate range, minimum, maximum, mean, median, std and the firrst (25%) and third (75%) quartiles for numeric variables.
range(iris$Petal.Width)
min(iris$Petal.Width)
max(iris$Petal.Width)
mean(iris$Petal.Width)
median(iris$Petal.Width)
sd(iris$Petal.Width)
quantile(iris$Petal.Width)
```
```{r}
## Calculate Variance
var(iris$Petal.Width)
```
```{r}
## Plot Petal.Width only
plot(iris$Petal.Width)
```


- Calculate Covariance and Correlation for Petal.Length  vs Petal.Width
```{r}
## Covariance
cov(iris$Petal.Length, iris$Petal.Width)
## Correlation
cor(iris$Petal.Length, iris$Petal.Width)
```
- Plot Petal.Length  vs Petal.Width
```{r}
plot(iris$Petal.Length, iris$Petal.Width)
```