---
title: "task1"
author: "Anna Palmqvist Sjövall"
date: "11 oktober 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Develop a Discriminant Analysis mode
## Use Iris dataset in R.
```{r}
library(datasets)
head(iris)
```

## Use sepal length and sepal width columns ONLY as 2 independent variables (X1 and X2) and last column as the response variable Y.
```{r}
data <- data.frame(X1 = iris$Sepal.Length, X2 = iris$Sepal.Width, Y = iris$Species)
```

## The dataset set has three categories (setosa, versicolor and virginica). Use setosa and versicolor ONLY. You need to REMOVE virginica rows.
```{r}
data <- subset(data, data$Y != "virginica")
```

## Discriminant Analysis
```{r}
#Split data into two classes- setosa and versicolor
c1<- data[data$Y=='setosa',]
c2<- data[data$Y=='versicolor',]

#Calculate the mean (average) vectors, covariance matrices and class probabilities. (X is Sepal length and width, Y is Species).
#c1
n1 <- length(c1$Y)
c1_P <- n1/length(data$Y)
c1_uX1 <- mean(c1$X1)
c1_uX2 <- mean(c1$X2)
c1_cov <- cov(c1[, 1:2])

#c2
n2 <- length(c2$Y)
c2_P <- n2/length(data$Y)
c2_uX1 <- mean(c2$X1)
c2_uX2 <- mean(c2$X2)
c2_cov <- cov(c2[, 1:2])

#Calculate the pooled covariance matrix and the coefficients of the model.
c <- (1 / (n1 + n2))*((n1*c1_cov) + (n2*c2_cov))
b <- solve(c)%*%(c(c1_uX1, c1_uX2) - c(c2_uX1, c2_uX2))
z <- (b[1,1]*data$X1) + (b[2,1]*data$X2)
data$Z <- z

#Calculate Z0
z0 <- (b[1,1]*((c1_uX1 + c2_uX1)/2)) + (b[2,1]*((c1_uX2 + c2_uX2)/2))
l <- -log(c1_P/c2_P)
data$Y_est <-ifelse(z-z0 > l, 'setosa','versicolor')
data['Z-Z0'] <- z - z0
```


## Display results and Print the final Accuracy.
```{r}
#results
data

#accuracy= (No. of correct / Total)*100. 
correct <- sum(data$Y == data$Y_est)
accuracy <- (correct / length(data$Y))*100
accuracy
```

