---
title: "Lab 6 - Part 2"
author: "Anna Palmqvist Sjövall"
date: "6 september 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2. Load college-admissions from you working directory (folder).
```{r}
# upload data
data <- read.csv("college-admissions.csv", header=TRUE, stringsAsFactors=FALSE)
head(data)
```

# 3. Normalise data X1 and X2 using Min Max method.
```{r}
data$X1 <- (data$X1 - min(data$X1))/(max(data$X1) - min(data$X1))
data$X2 <- (data$X2 - min(data$X2))/(max(data$X2) - min(data$X2))
head(data)
```

# 4. Apply Logistic Regression using b0=-0.0375, b1=-0.104290635, b2=-0.09564513761.
```{r}
b0 <- -0.0375
b1 <- -0.104290635
b2 <- -0.09564513761
data$z <- b0 + b1*data$X1 + b2*data$X2
data$Y_est <- 1/(1 + exp(-data$z))
data
```

# 5. Convert Yest into "0" and "1" as follows: if Yest <0.5 Yest =0 Else Yest=1
```{r}
data$Y_est <- ifelse(data$Y_est<0.5, 0, 1)
data
```

# 6. Calculate accuracy: Accuracy = (correct predictions / No. predictions made) * 100.
```{r}
correct <- sum(data$Y == data$Y_est)
accuracy <- (correct / length(data$Y))*100
accuracy
```

# 7. Apply above GDA steps with different update rules as follows: 
```{r}
theta <- 0.01
N <- length(data$Y)

for (i in 1:300) {
  SSE_b0 <- sum(data$Y_est - data$Y)/N
  SSE_b1 <- sum((data$Y_est - data$Y)*data$X1)/N
  SSE_b2 <- sum((data$Y_est - data$Y)*data$X2)/N
  b0 = b0 - (theta*SSE_b0)
  b1 = b1 - (theta*SSE_b1)
  b2 = b2 - (theta*SSE_b2)
  
  #Calculate z
  data$z = b0 + (b1*data$X1) + (b2*data$X2)
  #Calculate new Y_est
  data$Y_est <- 1/(1+exp(-data$z))
}
data$Y_est <- ifelse(data$Y_est<0.5, 0, 1)
data
```

# 8. Calculate accuracy: Accuracy = (correct predictions / No predictions made) * 100.
```{r}
correct <- sum(data$Y == data$Y_est)
accuracy <- (correct / length(data$Y))*100
accuracy
```
