---
title: "Lab 8"
author: "Anna Palmqvist Sjövall"
date: "17 september 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Use iris dataset in R
```{r}
library(datasets)
data(iris)
head(iris)
```

# 2. From iris dataset use only two columns: Petal.Length and Petal.Width
```{r}
d <- data.frame(pl = iris$Petal.Length, pw = iris$Petal.Width)
head(d)
```

# K-means Steps:

## Step 1. Set c1, c2, c3.
```{r}
c1= c(1.6, 0.2)
c2= c(4.6, 1.5)
c3= c(4.8, 1.8)
```

##Step 2. Each point is then assigned to the closest centroid based on the Euclidean distance.
```{r}
#Returns euclidean between two vectors
eucl_dist <- function(c, x) {
  sqrt(sum((c-x)^2))
}

#Returns matrix of distances between point and centroid
distance_matrix <- function(df, c) {
  distances <- matrix(NA, nrow=dim(df)[1], ncol=dim(c)[1])
  for(i in 1:dim(c)[1]) {
    distances[,i] <- apply(df, 1, eucl_dist, c[i,])
  }
  distances
}

centroid <- rbind(c1, c2, c3)
dist <- distance_matrix(d, centroid)
head(dist)
cluster <- apply(dist, 1, which.min)
cluster

```


## Step 3. The centroid of each cluster is then updated based on the points assigned to the cluster.
- Calculate the average of all points in each cluster to generate new centroid point.
```{r}
prev <- centroid
centroid <- apply(d, 2, tapply, cluster, mean)
centroid
```

## Step 4. We repeat Step 2 (assignment) and Step 3 (update) until no point changes clusters, or equivalently, until the centroids remain the same.
```{r}
while(any(prev != centroid)) {
  prev <- centroid

  dist <- distance_matrix(d, prev)
  cluster <- apply(dist, 1, which.min)
  centroid <- apply(d, 2, tapply, cluster, mean)
}
centroid
```

# Results
```{r}
d$cluster <- cluster
d$species <- iris$Species
d
```
