---
title: "Lab 4 - Dimensionality Reduction"
author: "Anna Palmqvist Sjövall"
date: "23 augusti 2018"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **CSE5DEV**

This lab coveres Lecture 4 - Applying PCA steps on iris dataset.



## Load the iris dataset

```{r}
# The datasets package needs to be loaded to access our data 
# For a full list of these datasets, type library(help = "datasets")
library(datasets)
data(iris)
head(iris)
```

## Make a copy of the data set: copy_iris
```{r}
copy_iris <- cbind(iris)
#Check so it doesn't reference the same memory space
tracemem(copy_iris)==tracemem(iris)
head(copy_iris)
```

## Remove last column (label feature or label class) from the dataset
```{r}
copy_iris <- copy_iris[, !((names(copy_iris) %in% "Species"))]
head(copy_iris)
```

## Normalise the dataset: X=x-mean(x) /Sd(x)
```{r}
normalize <- function(x) {
  return ((x-mean(x))/sd(x))
}
copy_iris <- as.data.frame(lapply(copy_iris, normalize))
head(copy_iris)
```

## Calculate the covariance matrix of the normalise the dataset: In R use Cor
```{r}
covariance_matrix <- cor(copy_iris)
covariance_matrix
```

## Find out the eigenvectors and eigenvalues using the covariance matrix:
```{r}
e <- eigen(covariance_matrix)
eigen_value <- e$values
eigen_vector <- e$vectors

eigen_value
eigen_vector
```

## Sort eigenvalues in descending order.
```{r}
#eigen() does this
```

## Calculate the Explained Variance = eigenvalue/sum(eigenvalues)
```{r}
exp_var <- eigen_value / sum(eigen_value)
exp_var
```

## Calculate the Cumulative Variance= Explained Variance (PCA1), Explained Variance (PCA1+PCA2), Explained Variance (PCA1+PCA2+PCA3), Explained Variance (PCA1+PCA2+PCA3+PCA4).
```{r}
cum_var <- c(exp_var[1], sum(exp_var[c(1,2)]), sum(exp_var[-4]), sum(exp_var)) 
cum_var
```

## Choose the best eigenvectors (the best two vectors) that correspond to the best largest eigenvalues (the best two eigenvalues). The best two that can lead to 90% or more based on the Cumulative Variance.
```{r}
best_eigen <- eigen_vector[, c(1,2)]
best_eigen
```

## Construct the projection matrix W from the selected two eigenvectors: W=DotProduct (Normalise_dataset, selected two eigenvectors) - in R: W=Normalise_dataset %*% selected eigenvectors.
```{r}
dim(copy_iris)
dim(best_eigen)
W <- data.matrix(copy_iris) %*% best_eigen
```

## Form the new data set in reduced dimensions- Added class label into W.
```{r}
W <- data.frame(W)
head(W)
species <- iris[, "Species"]
W$Species <- species
W
```